---
title: "COVID-19 USA"
author: "Prashant Sharma"
date: "30/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

We will see how is the impact of COvid-19 on the USA.
lets load our libraries first

```{r,message=FALSE}
library(plotly)
library(lubridate)
library(tidyr)
library(dplyr)
```

Our dataset for state-wise covid infection data.
```{r}
df = read.csv("data/us_states_covid19_daily.csv")

```
we need state and positive number of cases , hospitalized, totaltestsviral.

```{r}
req_cols = c("date","state","positive","hospitalized","totalTestsViral")
df_pos = df[,req_cols]
df_pos[is.na(df_pos)] <- 0
print((paste("number of NA rows:",sum(is.na(df_pos)))))
```
converting date into "date" datatype.
```{r}
df_pos$date = as.character(df_pos$date)
df_pos$date = as.Date(df_pos$date,format = ("%Y%m%d") , origin = "20200122")
class(df_pos$date)

```

```{r}
plot_ly(x = df_pos$date,y = df_pos$positive, color = factor(df_pos$state),
        mode = "lines") %>% layout(title = "USA COVID-19 trends from JAN to DEC 2020")
```


since our plot is looking messy we can try to find out which states are not doing well.

```{r}
df_pos$state = as.factor(df_pos$state)
class(df_pos$state)
df_state = group_by(df_pos , state)

```

```{r}
new_df = data.frame(summarise(df_state, Total_positive_cases = sum(positive) , 
                              Total_Hospitalized = sum(hospitalized), Total_testing = sum(totalTestsViral)))
head(new_df)
```
```{r}
sorted_df = new_df[order(new_df$Total_positive_cases,decreasing = TRUE),]
head(sorted_df) 
```

Top 10 states with highest number of positive cases.

```{r}
top_10 = sorted_df[1:10,]
head(top_10)
```

```{r}
state_name = as.character(top_10$state)
df_pos$state = as.character(df_pos$state)
new_df_pos = df_pos[df_pos$state %in% state_name,]
```

```{r}
plot_ly(x = new_df_pos$date,y = new_df_pos$positive, color = factor(new_df_pos$state),
        mode = "lines") %>% layout(title = "Top 10 US states COVID-19 trends from JAN to DEC 2020")
```
choropleth map of USA depicting Positivity rate.

```{r}
state_pop <- data.frame(State = new_df$state ,Positive_case = new_df$Total_positive_cases)

state_pop$hover <- with(state_pop, paste(State, "<br>", "positive_cases:", Positive_case))

borders <- list(color = toRGB("red"))

map_options <- list(
  scope = "usa",
  projection = list(type= "albers usa"),
  showlakes = TRUE,
  lakecolor = toRGB("white")
)
plot_ly(state_pop , z = state_pop$Positive_case , text = state_pop$hover ,locations = state_pop$State,
        type = "choropleth" ,locationmode = "USA-states",
        color = state_pop$Positive_case,colorscale = "Reds" , marker = list(line = borders)) %>%
  layout(title = "USA COVID-19 positivity" , geo = map_options)

```






